The language of generalization: Probability, vagueness, and interaction

### 2018-10-25 paper/appendix-bda.Rmd
update model results and analyses
revisions to appendices

In our three case studies, we compare an information-theoretic, computational model of endorsement to human endorsements of generalizations in language. 
The model has a single free parameter: the optimality parameter $\lambda$ in Eq. \ref{eq:S1}.
We analyze this model using a Bayesian data analytic approach, where we jointly infer this single model parameter $\lambda$ together with parameters that govern the prevalence priors $P(p)$ (Eq. \ref{eq:L0}) and referent prevalence $p$ (Eq. \ref{eq:S1}) for each item.
To pin down the prevalence prior and referent prevalence parameters, we use additional data directly related to those parameters (e.g., prior elicitation data).
Incorporating all data sources into a single Bayesian data analysis model is the appropriate way to track measurement uncertainty for all measurements.
<!-- Other model parameters (governing the prevalence prior $P(p_{fk})$ and the referent-prevalence $P_{fk'}$) are estimated using additional behavioral data (e.g., prior elicitation tasks).  -->
<!-- Througout the paper, we model all sources of data (prevalence prior data, endorsement data) using a single joint-inference Bayesian data analytic model. -->
In this appendix, we describe this procedure in more detail for each case study.

@@

In Case Study 1: Generic Language (Expt. 1b), we elicited the prevalence prior by asking participants about the prevalence of features for individual categories.
We performed an analogous elicitation in Case Study 3: Causal Language (Expt. 3a).
We describe the analysis using generics as our running example, but a parallel analysis was done for causals.

Participants' responses in the prior elicitation task can be thought of as *samples* from the prevalence prior distribution. 
Formally, we assume the prior data (analyzed independently for each property) was generated from one of two distributions: a distribution corresponding to those kinds with a stable causal mechanism that *could* give rise to the property ($\mathcal{D}_{stable}$) and a "transient cause" distribution corresponding to those kinds without a stable mechanism ($\mathcal{D}_{transient}$).
The "transient" distribution intuitively corresponds to categories that do not have the feature normally, but potentially could be caused by accidental forces (e.g., a lion, who through some genetic mutation, reproduces by laying eggs).
We model this distribution as a Beta distribuition that heavily favors probabilities near 0: $\text{Beta}(\gamma = 0.01, \delta = 100)$.^[
  Note that we use the noncanonical mean $\gamma$ and concentration $\xi$ (or, inverse-variance) parameterization of the Beta distribution rather than the canonical shape (or pseudocount) parameterization for ease of posterior inference. The shape parameterization can be recovered using: $\alpha = \gamma \cdot \xi; \beta = (1 - \gamma) \cdot \xi$.
]
The "stable" distribution is modeled as a Beta distribution with unknown parameters $\text{Beta}(\gamma, \xi)$.^[
Because the Beta distribution is not defined at the points 0 and 1, we add $\epsilon$ to the 0 responses and round 1 to 0.99.
Similar results can be obtained by rounding 0 to 0.01. 
Alternatively, the "transient" distribution could be defined as a Delta distribution at 0, and 0 responses could remain in their raw form.
Adjusting 1 to $1- \epsilon$ leads to improper inferences for this simple 2-component model, as $1 - \epsilon$ is only likely under a highly left-skewed distribution; treating 1 as $1- \epsilon$ disproportionately influences the shape of $\mathcal{D}_{stable}$, forcing it to favor probabilities close to 1.
This problem does not appear for 0 being adjusted to $\epsilon$ because the "transient" distribution already expects such low values.
]
Finally, we assume that these two components combine with mixture weighting $\phi$ such that the data we observe is $$P(d) = \phi\cdot \text{Beta} (d \mid \gamma, \xi) + (1 -  \phi) \cdot \text{Beta}(d \mid \gamma = 0.01, \xi = 100) $$.
We put the following priors over the latent parameters of the model:
\begin{eqnarray*}
\phi_i & \sim & \text{Uniform}(0, 1) \\
\gamma_i & \sim & \text{Uniform}(0, 1) \\
\xi_i & \sim & \text{Uniform}(0, 100)
\end{eqnarray*}
where $i$ ranges over the different properties (e.g., *lays eggs*, *carries malaria*).

@@

In Case Study 2 (Expt. 2a), we asked participants about parameters of this mixture model (by having participants answer questions about different kinds of people) rather than having participants give samples (e.g., by listing their friends and family members, and rating how often they did certain actions).
In pilot testing, we found these different methodologies to give comparable results and we opted to ask about hypothetical people to probe about potentially undesirable habits of participants' friends and familiy (e.g., how often they use cocaine).
The questions used in this structured elicitation task are  described in the main text.

@@

In Case Study 1, we used participants prevalence ratings for the category-of-interest in our generic sentences as the referent-prevalence that is used in the endorsement model (Eq. \ref{eq:S1}).
For a given generic sentence (e.g., "Robins lay eggs"), we took the prevalence ratings for the referent-category (e.g., the percentage of robins that lay eggs) from the prior elicitation task (Expt. 1b) and assumed those were generated from a single Beta distribution.
We assumed the following priors on the parameters:
\begin{eqnarray*}
\gamma_i & \sim & \text{Uniform}(0, 1) \\
\xi_i & \sim & \text{Uniform}(0, 100)
\end{eqnarray*}
We took samples from the posterior predictive of this Beta distribution (i.e., reconstructed prevalence ratings) as the *referent-prevalence* used in the model.

@@

To fit the generic endorsement models, we incorporate them into the Bayesian data analytic model of the prevalence prior data (described above) to create a single, joint-inference model where the optimality parameter $\lambda$ (Eq. \ref{eq:S1}) is inferred jointly with all the other latent parameters of the full model (the referent-prevalence $p$ for each category $k$ and property $f$ and the parameters of the prevalence priors $P(p)$ for each property $f$) using data from Expt. 1a \& b (Figure \ref{fig:genericsModelDiagram}).
For the parameters of the prevalence priors, we use the same priors described in Expt. 1b; for the speaker optimality parameter, we use a prior with a range consistent with previous literature using the same model class: $\lambda \sim \text{Uniform}(0,5)$.
We learn about the \emph{a posteriori} credible values of the joint inference models by collecting samples from 3 MCMC chains of 150,000 iterations removing the first 50,000 iterations for burn-in, using an incrementalized version of the Metropolis-Hastings algorithm [@Ritchie2016]. This algorithm is useful for models with many variables that only affect a subset of the full model's predictions (e.g., models with by-item or by-participant parameters, wherein those additional parameters only influence predictions for those items or participants).

```{r genericsModelDiagram, fig.cap="Quasi-graphical model corresponding to the fully Bayesian data analysis of the endorsement model. The prevalence prior data $d^{pr}_{f}$ is assumed to be generated from the mixture model validated in Expt. 1a. The referent-prevalence $d^{pr}_{k,f}$ is generated from a latent prevalence $p_{fk}$. The fixed-threshold (control model) or uncertain-threshold model take the prevalence priors $P(p)$, the referent-prevalence $p$, and a single speaker optimality parameter $\\lambda$ as input. This overall structured is repeated (except $\\lambda$) for each of the unique properties $f$ and categories $k$ that correspond to the generic sentences in our stimulus set. Note that S corresponds to a probabilistic *function* and not a *random variable* that is standard in graphical model notation; S cannot be represented by a graphical model because it has recursion. "}
knitr::include_graphics("figs/generics_bayesnetmodel.pdf", dpi = 108)
```

@@

In addition to examining the posterior predictive distribution on endorsement judgments (presented in main text), we examined the marginal posteriors on parameters of the prevalence priors and referent-prevalence.
These marginal distributions are important to confirm that they have not changed substantially from the parameters inferred from their primary data sources in isolation.
For example, when modeling the referent-prevalence data in isolation, the model infers that roughly 65\% of robins lay eggs, as that is what participants on average produce in the prevalence elicitation task.^[
  Most participants report that 50\% of robins lay eggs, while a minority respond 100\%.
]
If the joint inference model (which models all data sources---referent prevalence, prevalence prior, and generic endorsement---simultaneously) infers referent-prevalence values substantially different from those inferred by a model of referent prevalence in isolation, that would suggest that the joint-model is distorting the prevalence parameters to accomodate the endorsement data.
Such a result would call into question the inferences we as scientists derive from the joint inference model.
For example, incorporating a linear regression model (of the kind presented as alterantive models in the main text) into this Bayesian joint-inference analysis model produces posterior predictions that match the generic endorsement data surprisingly well (e.g., that model predicts "Robins lay eggs" is true).
Such a model is only able to do this, however, by distorting the referent-prevalence data, inferring that 100\% of robins lay eggs; thus, the linear model in this joint-inference analysis framework sacrifices its goodness-of-fit to the referent-prevalence data in order to increase its goodness-of-fit to the endorsement data. ^[
  This distortion effect is why we accout for measurement uncertainty in the linear models by bootstrapping the data that forms their predictors, rather than performing a Bayesian analysis.
]
Such a distortion manifests as a difference between the inferred parameters given only the referent-prevalence data and given the full joint model (all data sources simultaneously).

>>>>

In our three case studies, we compare an information-theoretic, computational model of endorsement to human endorsements of generalizations in language. 
The model has a single free parameter: the optimality parameter $\lambda$ in Eq. \ref{eq:S1}.
We analyze this model using a Bayesian data analytic approach, where we jointly infer this single model parameter $\lambda$ together with parameters that govern the prevalence priors $P(p)$ (Eq. \ref{eq:L0}) and referent prevalence $p$ (Eq. \ref{eq:S1}) for each item.
To pin down the prevalence prior and referent prevalence parameters, we use additional data directly related to those parameters (e.g., prior elicitation data).
Incorporating all data sources into a single Bayesian data analysis model is the appropriate way to track measurement uncertainty for all measurements simultaneously.
<!-- Other model parameters (governing the prevalence prior $P(p_{fk})$ and the referent-prevalence $P_{fk'}$) are estimated using additional behavioral data (e.g., prior elicitation tasks).  -->
<!-- Througout the paper, we model all sources of data (prevalence prior data, endorsement data) using a single joint-inference Bayesian data analytic model. -->
In this appendix, we describe this procedure in more detail for each case study.

@@

In Case Study 1: Generic Language (Expt. 1b), we elicited the prevalence prior by asking participants about the prevalence of features for individual categories.
We performed an analogous elicitation in Case Study 3: Causal Language (Expt. 3a).
We describe the analysis using generics as our running example, but a parallel analysis was done for causal language.

Participants' responses in the prior elicitation task can be thought of as samples from the prevalence prior distribution. 
Formally, we assume the prior data (analyzed independently for each property) was generated from one of two underlying distributions: a distribution corresponding to those kinds with a stable causal mechanism that *could* give rise to the property ($\mathcal{D}_{stable}$) and a "transient cause" distribution corresponding to those kinds without a stable mechanism ($\mathcal{D}_{transient}$).
The "transient" distribution intuitively corresponds to categories that do not have the feature normally, but potentially could acquire the feature by accidental forces (e.g., a lion, who through some genetic mutation, reproduces by laying eggs).
We model this distribution as a Beta distribuition that heavily favors probabilities near 0: $\text{Beta}(\gamma = 0.01, \delta = 100)$.^[
  Note that we use the noncanonical mean $\gamma$ and concentration $\xi$ (or, inverse-variance) parameterization of the Beta distribution rather than the canonical shape (or pseudocount) parameterization for ease of posterior inference. The shape parameterization can be recovered using: $\alpha = \gamma \cdot \xi; \beta = (1 - \gamma) \cdot \xi$.
]
The "stable" distribution is modeled as a Beta distribution with unknown parameters $\text{Beta}(\gamma, \xi)$.^[
Because the Beta distribution is not defined at the points 0 and 1, we add $\epsilon$ to the 0 responses and round 1 to 0.99.
Similar results can be obtained by rounding 0 to 0.01. 
Alternatively, the "transient" distribution could be defined as a Delta distribution at 0, and 0 responses could remain in their raw form.
Adjusting 1 to $1- \epsilon$ leads to improper inferences for this simple 2-component model, as $1 - \epsilon$ is only likely under a highly left-skewed distribution; treating 1 as $1- \epsilon$ disproportionately influences the shape of $\mathcal{D}_{stable}$, forcing it to favor probabilities close to 1.
This problem does not appear for 0 being adjusted to $\epsilon$ because the "transient" distribution already expects such low values.
]
Finally, we assume that these two components combine with mixture weighting $\phi$ such that the data we observe is $$P(d) = \phi\cdot \text{Beta} (d \mid \gamma, \xi) + (1 -  \phi) \cdot \text{Beta}(d \mid \gamma = 0.01, \xi = 100) $$.
We put the following priors over the latent parameters of the model:
\begin{eqnarray*}
\phi_i & \sim & \text{Uniform}(0, 1) \\
\gamma_i & \sim & \text{Uniform}(0, 1) \\
\xi_i & \sim & \text{Uniform}(0, 100)
\end{eqnarray*}
where $i$ ranges over the different properties (e.g., *lays eggs*, *carries malaria*).

@@

In Case Study 2 on habitual language (Expt. 2a), we asked participants about parameters of this mixture model (by having participants answer questions about different kinds of people) rather than having participants give samples (e.g., by listing their friends and family members, and rating how often they did certain actions).
In pilot testing, we found these different methodologies to give comparable results and we opted to ask about hypothetical people to probe about potentially undesirable habits of participants' friends and familiy (e.g., how often they use cocaine).
The questions used in this structured elicitation task are  described in the main text.

@@

In Case Study 1 (generics), we used participants prevalence ratings for the category-of-interest in our generic sentences as the referent-prevalence that is used in the endorsement model (Eq. \ref{eq:S1}).
For a given generic sentence (e.g., "Robins lay eggs"), we took the prevalence ratings for the referent-category (e.g., the percentage of robins that lay eggs) from the prior elicitation task (Expt. 1b) and assumed those were generated from a single Beta distribution.
We assumed the following priors on the parameters:
\begin{eqnarray*}
\gamma_i & \sim & \text{Uniform}(0, 1) \\
\xi_i & \sim & \text{Uniform}(0, 100)
\end{eqnarray*}
We took samples from the posterior predictive of this Beta distribution (i.e., reconstructed prevalence ratings) as the *referent-prevalence* used in the model.

@@

To fit the generic endorsement models, we incorporate them into the Bayesian data analytic model of the prevalence prior data (described above) to create a single, joint-inference model where the optimality parameter $\lambda$ (Eq. \ref{eq:S1}) is inferred jointly with all the other latent parameters of the full model (the referent-prevalence $p$ for each category $k$ and property $f$ and the parameters of the prevalence priors $P(p)$ for each property $f$) using data from Expt. 1a \& b (Figure \ref{fig:genericsModelDiagram}).
For the parameters of the prevalence priors, we use the same priors described in Expt. 1b; for the speaker optimality parameter, we use a prior with a range consistent with the previous literature that uses the same model class: $\lambda \sim \text{Uniform}(0,5)$.
We learn about the \emph{a posteriori} credible values of the joint inference models by collecting samples from 3 MCMC chains of 150,000 iterations removing the first 50,000 iterations for burn-in, using an incrementalized version of the Metropolis-Hastings algorithm [@Ritchie2016]. This algorithm is useful for models with many variables that only affect a subset of the full model's predictions (e.g., models with by-item or by-participant parameters, wherein those additional parameters only influence predictions for those items or participants).

```{r genericsModelDiagram, fig.cap="Quasi-graphical model corresponding to the fully Bayesian data analysis of the endorsement model for Case Study 1 (generic language). The prevalence prior data $d^{pr}_{f}$ is assumed to be generated from the mixture model validated in Expt. 1a, which has three parameters: mean of the stable-cause distribution $\\gamma_{f}$ and concentration (inverse-variance) of the stable-cause distribution $\\xi{f}$ and the mixture parameter $\\phi_{f}$. The referent-prevalence $d^{pr}_{k,f}$ is generated from a Beta distribution with paramters: mean $\\gamma_{k,f}$ and concentration (inverse-variance) $\\xi{k,f}$. The posterior predictives of the prevalence prior $P(p_{fk'})$ and the referent prevalence $p_{fk}$ are then fed into the RSA speaker model $S$, which predicts the generic endorsement data $d^{end}_{k, f}$. The speaker model $S$ also takes in the single free parameter $\\lambda$ which operates as a soft-max function. This overall structured is repeated (except $\\lambda$) for each of the unique properties $f$ and categories $k$ that correspond to the generic sentences in our stimulus set. Note that $S$ corresponds to a probabilistic function and not a random variable that is standard in graphical model notation; $S$ cannot be represented by a graphical model because it has recursion. This entire BDA model is duplicated for the lesioned, fixed-threshold model (which only differs in the definition of $S$). The BDA model for habituals and causals mirrors this one, except they do not infer a referent prevalence $p_{fk}$ (they are assumed to be the same as those experimentally supplied to participants).", out.width = "0.8\\textwidth", fig.align="center"}
knitr::include_graphics("figs/generics_bayesnetmodel.pdf", dpi = 108)
```

@@

In addition to examining the posterior predictive distribution on endorsement judgments (presented in main text), we examined the marginal posteriors on parameters of the prevalence priors and referent-prevalence.
These marginal distributions are important to confirm that they have not changed substantially from the parameters inferred from their primary data sources in isolation.
For example, when modeling the referent-prevalence data in isolation, the model infers that roughly 65\% of robins lay eggs, as that is what participants on average produce in the prevalence elicitation task.^[
  Most participants report that 50\% of robins lay eggs, while a minority respond 100\%.
]
If the joint inference model (which models all data sources---referent prevalence, prevalence prior, and generic endorsement---simultaneously) infers referent-prevalence values substantially different from those inferred by a model of referent prevalence in isolation, that would suggest that the joint-model is distorting the prevalence parameters to accommodate the endorsement data.
Such a result would call into question the inferences we as scientists derive from the joint inference model.
For example, incorporating a linear regression model (of the kind presented as alterantive models in the main text) into this Bayesian joint-inference analysis model produces posterior predictions that match the generic endorsement data surprisingly well (e.g., that model predicts "Robins lay eggs" is true).
Such a model is only able to do this, however, by distorting the referent-prevalence data, inferring that 100\% of robins lay eggs; thus, the linear model in this joint-inference analysis framework sacrifices its goodness-of-fit to the referent-prevalence data in order to increase its goodness-of-fit to the endorsement data. ^[
  This distortion effect is why we accout for measurement uncertainty in the linear models by bootstrapping the data that forms their predictors, rather than performing a Bayesian analysis.
]
Such a distortion manifests as a difference between the inferred parameters given only the referent-prevalence data and given the full joint model (all data sources simultaneously).

### 2018-10-25 paper/appendix-cueValidity.Rmd
revisions to appendices

In Experiment 1, we articulated an alternative model by measuring cue validity (and prevalence) and predicting generic endorsement from a linear combination of these parameters.
In a small review of the literature, we discovered different methods for measuring cue validity; in piloting, these different methods lead to different results. 
Therefore, we propose three *a priori* desiderata that a measurement of cue validity should satisfy.
We describe two experiments that represent the primary methods for measuring cue validity and compare them with these desiderata in mind.
Finally, we compare the cue validity measured using these different methods to cue validity derived from our prevalence prior elicitation task (Expt. 1b, main text).

@@

Measuring cue validity involves collecting participants' judgments that relate to the probability that an exemplar is a member of a kind given that it has a feature : $P(x \in k \mid x \in f)$.
There are several ways one could measure cue validity. 
Here we consider two measures: One has participants estimate the probability $P(x \in k \mid x \in f)$ directly, a common technique in the literature on generic language [e.g., @Khemlani2012; @Prasada2013], and another has participants freely produce categories given a feature [i.e., draw a sample from the conditional distribution on kinds given a feature; @Cree2006].
We will refer to the former the *direct question* method and the latter the *free production* method.

Are the *direct question* and the *free productions* equally valid for measuring cue validity?
We propose *a priori* three boundary conditions that the measurement should be able to satisfy.
For each case, we provide four examples from our larger stimulus set on generics (Case Study 1) which will be used to evaluate each measure.

\begin{enumerate}
\item{Completely diagnostic features: Some features are only present in one (or a very small number) of categories. Examples include: \emph{carrying malaria} (mosquitos), \emph{carrying Lyme disease} (ticks, deer), \emph{having manes} (lions), \emph{having pouches} (marsupials, including most famously: kangaroos). The cue validity of these features for the corresponding categories should be very high (at least 0.5 and possibly close to 1).}
\item{Completely absent features: Many features are completely absent in many kinds. For these, the cue validity should be extremely low or 0. There are infinite examples. The ones we will use are \emph{has wings} (leopard), \emph{has a mane} (shark), \emph{has spots} (kangaroo), \emph{has a pouch} (tiger). }
\item{Completely undiagnostic features: A number of features are shared by almost every category. The ones we will use are: \emph{is female} (robin), \emph{is male} (lion), \emph{is juvenile} (kangaroo), \emph{is full-grown} (leopard). The cue validity of these features for particular categories should be extremely low or 0. Learning that an entity is female tells you almost nothing about what kind of animal it is.}
\end{enumerate}

We collected cue validity ratings by running both a direct question and a free production experiment.
For the free production experiment, the cue validity is the proportion of responses of the target category (e.g., "mosquitos") for the property (e.g., "carries malaria").
Of primary interest is the measurement for the desiderata items described above.
Links to the experiments can be found on \url{https://mhtess.github.io}.

@@

To process the free production, we forced all characters in a repsonse to lower case, removed spaces, and made all terms into singular terms (e.g., "lions" --> "lion").
As well, "mosquito" was a commonly mispelled label; we counted anything that started with "mosqu", "mesqu", "misqu", "mosiq" as "mosquito".

@@

```{r cvBothquestionsBarplots, fig.width=10, fig.asp=0.7, fig.cap="A: Empirically measured cue validity for two different tasks. Items are grouped by whether the property is never present in the category (false), the property is always present in the category and every other category (undiagnostic), or present in the category and absent from most other categories (diagnostic). Dotted lines denote theoretical cue validity representing the desiderata (see text).  B: Raw empirical distributions for direct question task. C: Correspondence between measured cue validity and prevalence prior derived cue validity.", out.width="\\textwidth", fig.pos="!h", cache = F}
load("cached_results/appendix_cueValidity_results.RData")
# d.cv.bothtasks, d.cv.targets, undiagnostic.props
# md.fp, md.dq

@@

The points of largest deviation for the free-production measurement from the prevalence prior derived measurement occur where the prevalence prior derived measure rates the cue validity as relatively high when the free-production measure gives the item low cue validity (two black points in Figure\ \@ref(fig:cvBothquestionsBarplots)C, high X-value, low Y-value).
These two items are: ("is red", "cardinal") and ("is white", "swan").
These items should have relatively low cue validity and are overestimated by the prevalence prior because of the prior on categoryies $P(k)$ (i.e., because they were supplied to every participant and thus get a higher weighted in the prior for deriving cue validity).
In comparison, the direct question measurement consistently overestimates cue validity relative to the cue validity derived from the prevalence prior for almost all of the items.

@@

*Cue validity* is a commonly used measurement for understanding generic truth judgments [e.g., @Khemlani2012; @Prasada2013].
We observed different measurements used in the literature and articulated three *a priori* desiderata to validate a measure of cue validity.
We found that the "free production" measurement (i.e., participants freely produced categories given a feature), and not the direct question measurement (i.e., participants provide a likelihood judgment of a particular category given the feature), satisfied all three boundary conditions. 
In addition, cue validity derived from our prevalence prior measurement (Expt. 1b) also satisfied these boundary conditions. 
We conclude that researchers interested in comparing cue validity to generic truth judgments should use a free production paradigm for measuring cue validity.

>>>>

In Experiment 1, we articulated an alternative model by measuring cue validity (and prevalence) and predicting generic endorsement from a linear combination of these parameters.
In a small review of the literature, we discovered different methods for measuring cue validity; in piloting, we found these different methods led to different results. 
Therefore, we propose three *a priori* desiderata that a measurement of cue validity should satisfy.
We describe two experiments that represent the primary methods for measuring cue validity and compare them with these desiderata in mind.
Finally, we compare the cue validity measured using these different methods to cue validity derived from our prevalence prior elicitation task (Expt. 1b, main text).

@@

Measuring cue validity involves collecting participants' judgments that relate to the probability that an exemplar is a member of a kind given that it has a feature: $P(x \in k \mid x \in f)$.
There are several ways one could measure cue validity. 
Here we consider two measures: The first has participants estimate the cue validity probability $P(x \in k \mid x \in f)$ directly, a common technique in the literature on generic language [e.g., @Khemlani2012; @Prasada2013], and another has participants freely produce categories given a feature [i.e., draw a sample from the conditional distribution on kinds given a feature; @Cree2006].
We will refer to the former the *direct question* method and the latter the *free production* method.

Are the *direct question* and the *free productions* equally valid for measuring cue validity?
We propose *a priori* three boundary conditions that a measurement of cue validity should satisfy.
For each case, we provide four examples from our larger stimulus set on generics (Case Study 1) which will be used to evaluate each measure.

\begin{enumerate}
\item{Completely diagnostic features: Some features are only present in one (or a very small number) of categories. Examples include: \emph{carrying malaria} (mosquitos), \emph{carrying Lyme disease} (ticks, deer), \emph{having manes} (lions), \emph{having pouches} (marsupials, including most famously: kangaroos). The cue validity of these features for the corresponding categories should be very high (at least 0.5 and possibly close to 1).}
\item{Completely absent features: Many features are completely absent in many kinds. For these, the cue validity should be extremely low or 0. There are infinite examples. The ones we will use are \emph{has wings} (leopard), \emph{has a mane} (shark), \emph{has spots} (kangaroo), \emph{has a pouch} (tiger). }
\item{Completely undiagnostic features: A number of features are shared by almost every category. The cue validity of these features for particular categories should be extremely low or 0. The ones we will use are: \emph{is female} (robin), \emph{is male} (lion), \emph{is juvenile} (kangaroo), \emph{is full-grown} (leopard).  Learning that an entity is female tells you almost nothing about what kind of animal it is.}
\end{enumerate}

We collected cue validity ratings by running both a direct question and a free production experiment.
For the free production experiment, the cue validity is the proportion of responses of the target category (e.g., "mosquitos") for the property (e.g., "carries malaria").
Of primary interest is the measurement for the desiderata items described above.
Links to the experiments can be found on \url{https://github.com/mhtess/genlang-paper}.

@@

To process the free production, we forced all characters in a response to lower case, removed spaces, and made all terms into singular terms (e.g., "lions" --> "lion").
As well, "mosquito" was a commonly misspelled label; we counted anything that started with "mosqu", "mesqu", "misqu", "mosiq" as "mosquito".

@@

```{r cvBothquestionsBarplots, fig.width=10, fig.asp=0.7, fig.cap="A: Empirically measured cue validity for two different tasks. Items are grouped by whether the property is never present in the category (false), the property is always present in the category and every other category (undiagnostic), or present in the category and absent from most other categories (diagnostic). Dotted lines denote theoretical cue validity representing the desiderata (see text).  B: Raw empirical distributions for the undiagnostic features in the direct question task. C: Correspondence between measured cue validity and prevalence prior derived cue validity.", out.width="\\textwidth", fig.pos="!h", cache = F}
load("cached_results/appendix_cueValidity_results.RData")
# d.cv.bothtasks, d.cv.targets, undiagnostic.props
# md.fp, md.dq

@@

The points of largest deviation for the free-production measurement from the prevalence prior derived measurement occur where the prevalence prior derived measure rates the cue validity as relatively high when the free-production measure gives the item low cue validity (two black points in Figure\ \@ref(fig:cvBothquestionsBarplots)C, high X-value, low Y-value).
These two items are: ("is red", "cardinal") and ("is white", "swan").
These items should have relatively low cue validity and are overestimated by the prevalence prior because of the prior on categories $P(k)$ over-represents the categories that were supplied to every participant and thus get a higher weighted in the prior for deriving cue validity.^[
  This deviation could be reduced in future experiments by performing a fully free production version of the prevalence prior task (i.e., without supplying the referent categories).
]
In comparison, the direct question measurement overestimates cue validity for almost all of the items relative to the cue validity derived from the prevalence prior.

@@

*Cue validity* is a commonly used measurement for understanding generic truth judgments [e.g., @Khemlani2012; @Prasada2013].
We observed different measurements used in the literature and articulated three *a priori* desiderata to validate a measure of cue validity.
We found that the "free production" measurement (i.e., participants freely produced categories given a feature), and not the direct question measurement (i.e., participants provide a likelihood judgment of a particular category given the feature), satisfied all three boundary conditions. 
In addition, cue validity derived from our prevalence prior measurement (Expt. 1b) also satisfied these boundary conditions. 
Researchers interested in comparing cue validity to generic truth judgments should use a free production paradigm for measuring cue validity.

### 2018-10-25 paper/appendix-cvDerivation.Rmd
revisions to appendices

Cue validity is defined for a particular category--property pair (e.g., *mosquitos* and *carry malaria*), and relates to referent-prevalence (e.g., how many mosquitos carry malaria) via the prevalence of *carrying malaria* in other categories $k'$ by Bayes' Rule:
$$ P(k \mid f) = \frac{P(f \mid k) \cdot P(k)}{Z} $$
where $Z =  \sum_{k' \in K} P( f \mid k') \cdot P( k')$

The prevalence prior $P(p)$ (Eq. \ref{eq:L0}) is a probability distribution over prevalence for different categories $k'$.
<!-- For ease of exposition, let $x = p = P(f \mid k)$; that is, $P(x) = P(p)$. -->

>>>>

Cue validity is defined for a particular category--property pair (e.g., *mosquitos* and *carry malaria*), and relates to the referent prevalence (e.g., how many mosquitos carry malaria) via Bayes' Rule:
$$ P(k \mid f) = \frac{P(f \mid k) \cdot P(k)}{Z} $$
where $Z =  \sum_{k' \in K} P( f \mid k') \cdot P( k')$, the average or marginal prevalence of the feature (e.g., *carrying malaria*) in other categories $k'$.

The prevalence prior $P(p)$ using the generic interpretation model (Eq. \ref{eq:L0}) is a probability distribution over prevalence for different categories $k'$.
<!-- For ease of exposition, let $x = p = P(f \mid k)$; that is, $P(x) = P(p)$. -->
